Networked Character Devices

### 2011-12-13 report/paper.tex
paper: make a pdf index, make references clickable
Added final changes and fixed hyperref version dependency

\usepackage[text={6.5in, 9in}, centering]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}

@@

@@

We have designed and implemented Network Character Devices (NCD) to
enable a new approach to using and sharing devices. Today, networking 
technology is becoming faster and more reliable at incredible rates. 
More functionality that has been traditionally thought to only be 
feasibly locally accessible can now be pushed to remote
machines. Using client-server operation, our project aims to allow 
a character device (or devices) present on a server to be exported to
a client over a network connection. While character devices come in 
many forms, from Web cams, mice and monitors to USB flash drives and 
external hard drives, NCD works below where these devices are 
interpreted and exports their device file over IP. While designed for
character devices, this generality will actually allow us to export 
any file, which in a *nix system means almost anything on the
system. Being able to dynamically export devices over network
connections suggests a new cloud-like paradigm of device usage where 
locality is required only for reduced latency, which may not be a
concern for certain devices.

When a device is being exported by NCD we would like it to give the
appearance of having essentially ``moved'' the device to the client 
by forbidding user level accesses on the server. This is not part 
of our current iteration of NCD, but it will be included as higher
level functionality is implemented. Acquiring a remote device will be
done through an advertisement protocol that exists in user space so
that details may be supplied about the server and how the connection
should be made (e.g. UDP vs TCP).  After this NCD's functionality on
the client exists in kernel space so as to maintain transparency to
user level operations. Once imported, a device should be accessed just
like a normal local device.  This is achieved by the NCD kernel module
intercepting accesses and sending them out over the network to the
server. On the server, a userspace daemon accesses the physical
device to emulate the client's activity. The overall structure of the
NCD system is present in Figure \ref{fig:system-full}.

@@

Currently we have achieved a successful implementation of the low
level functionality for NCD. The client uses a kernel module written
for Linux 2.6.x+ while the server uses a userspace daemon. NCD is
still rough and in need of further development, as it tends to be
crash prone if not treated with care. The groundwork has been laid for
exporting any device, however many device subsystems access their
devices by avoiding the character device interface and these must be
dealt with on an individual basis. Overall, the path for future work
is very clear.

@@

NFS was the first implementation of truly general information sharing
over remote network connections.  However, NFS uses a specific file
system and cannot interface directly with devices \cite{nfs-source}.

NBD (Network Block Device) was included into the Linux kernel starting
with version 2.1.101. It enables users to export block devices over the
network, allowing use of any filesystem. However, NBD provides no
exclusive access guarantees and users must ascertain on their own
whether the remote drive is being used \cite{nbd-source}.

CUSE (Character Device in User Space) is an extension of FUSE that
allows bringing a character device into a FUSE filesystem.  This allows
users to access the device directly and potentially export it using
user-level networking \cite{cuse-article}.

USB/IP is a project to share USB devices over IP. This project is
very similar to NCD (as USB devices are character devices) but is
tailored specifically to allow sharing USB devices over a
network connection by intercepting USB specific data and encapsulating
it in IP \cite{usbip-source}.

Plan 9 is a distributed operating system developed as a successor to
Unix by Bell Labs.  It implicitly allows sharing devices over a
network via file interfaces. However, Plan 9 does not currently see
wide deployment \cite{plan9-pike}.

@@

The client is currently implemented entirely within an out-of-tree kernel
module. This is arguably not the most robust, nor secure, solution. See
section \ref{sec:futurework} for details on splitting out some functionality
into a user-space helper.

@@

The server is a typical user-space daemon. It
currently only services a single file. Whether or not the server should
handle multiple exports is still up for evaluation.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{proto-g.png}
  \caption{NCD operates over a simple request reply protocol. Bulk data
    immediately follows formal messages, as required.}
  \label{fig:protocol}
\end{figure}

The protocol is implemented as a simple request-reply system over
\texttt{INET} sockets. All request types are encapsulated within a
single request structure (see section \ref{sec:implementation}).
Similarly, there is only a single reply structure to account for all
reply types.

@@

Because there is only one type of packet for a request, and another for
a reply, (listing \ref{lst:protocol}) there is a small amount of wasted
space on some file operations. We have done our best to reduce this
overhead as much as possible, making use of a cascade of nested unions
and structs.

@@

The client resides entirely within a kernel module \footnote{The merits
of including everything within a kernel module are discussed in section
\ref{sec:currentstatus}.}. A typical userspace program can be sloppy
with cleanup because all memory is freed on termination. In contrast, a
kernel module must be very careful because its lifetime is shorter than
the kernel itself. Therefore it is very import to correctly handle
errors and cleanup all memory in the \texttt{module\_init} and
\texttt{module\_exit} routines.

Some of the basic features of the NCD kernel module are listed in
listing \ref{lst:client}. NCD dynamically requests a major and
minor number from the running kernel \footnote{Statically assigned
major/minor numbers are no longer required and are discouraged for new
modules.}. These numbers are stored in a \texttt{struct dev\_t}. Once
these numbers are assigned, the kernel will pass all file operations on
character devices with the NCD major/minor number to the NCD module.

@@

In the snippet given in listing \ref{lst:server}, a \texttt{read()} file
operation is handled. A buffer must be allocated with a size
indicated by the original request. Next, the server can perform the
corresponding operation and send a reply message to the client. Finally,
the bulk data is sent to the client. All buffering is performed by the
underlying \texttt{SOCK\_STREAM} and TCP protocols.

@@

A lot of magic is performed on character devices in userspace by
the Linux \texttt{udev} system.
Currently we are using \texttt{udev} to manage the
permissions of imported devices (listing \ref{lst:udevrule}).

@@

For more functionality, imported character devices must have certain
attributes such as those for the keyboard in listing \ref{lst:udevattr}.
Unfortunately setting them retroactively does not convince all programs
of a particular devices type. Therefore interfacing with specific kernel
subsystems such as \texttt{input} and \texttt{usb} will be required in
the future for enhanced functionality.

@@

This set of core functionality provides a working proof-of-concept
Network Character Devices system. This system fully supports the
export of a single character device from server to client over the
network and demonstrates the basic use and utility of such a system.

@@

Most modern Operating Systems, including Linux, assume that any device
physically connected to the system is available for the system to
use. As mentioned previously, Network Character Devices break down this
assumption. Thus, one of the challenges involved in implementing an NCD
system is to remove the OS's assumption that any connected devices in
available for local use. Modern-day plug-and-play auto-configuration
systems make this task even more difficult, and the means for breaking
down the availability assumption tend to be unique to a given
OS. Effectively removing this assumption is one of the larger,
still-open, NCD challenges.

@@

\begin{packed_item}
\item Add multi-device, multi-server support
\item Add kernel sub-system handler layer (HID, etc)
\item Add device protection and removal capabilities to server system
\item Add advertisement protocol
\item Separate networking and memory-intensive code from client kernel module
\end{packed_item}

The current NCD system only support a single device per server and per
client. To be truly useful and to enable the construction of the
kinds of system we envision NCD being used for (Section
\ref{sec:potentialuses}), we must at least
expand the client-side
system to support importing an arbitrary number of devices from one
or more servers. Ideally, servers should also be capable of serving
multiple devices, but this is of less importance than multi-device
client handling.

@@

@@

Finally, the current client-side NCD implementation performs both direct
network access and memory-intensive processing directly in the kernel
code. For security reasons, it is desirable to separate the direct
network communication from the kernel code to user-space code where
the repercussions of a security breach as less extreme.
For performance reasons, it is desirable to move memory
intensive operations from the kernel code (where they are limited to
the available kernel memory space) to user-space code where the memory
availability is far greater and the use restrictions more relaxed. To
solve these issues, we
foresee building a client-daemon that will handle the bulk to the
client-side operations and that will, in turn, communicate directly with
a more minimal client kernel module only when necessary.

@@

KVM (Keyboard-Video-Mouse) systems are commonly used in data centers
and other areas where a user wishes to use a single keyboard, video,
and mouse, to control multiple machines. KVM systems range from simple KVM
switches that allow the user to select which local machine is
physically connected to his or her i/o devices to
networked KVM systems that allow the user to use a
local KVM setup to interface with a computer located elsewhere on the
network. The latter class of networked KVM devices traditionally require
expensive propriety hardware.

@@

Another example of the novel systems that can be built use Network
Charter Devices is a distributed random number generator. Linux has a
fairly robust random number generator built into it and accessible as
a Character Device at \texttt{/dev/random}. This RNG relies on the
local entropy pool (generally filled by human interaction, interrupt
timings, and other high entropy events) to generate random
numbers. This, the speed at which random numbers can be generated is
limited on any single system to the locally available
entropy. This can severely hinder the
performance of systems that require random numbers\footnote{Linux also
provides the \texttt{/dev/urandom} RNG which is a pseudorandom number
generator seeded off the true
\texttt{/dev/random}. \texttt{/dev/urandom} trades ``randomness'' in
favor of speed. It is thus handy when one just needs to quickly
generate pseudorandom numbers, but is not appropriate for cryptography or
other secure random system.}.

@@

Figure \ref{fig:rng} shows how such a system might work. In this
distributed RNG system, multiple hosts act as NCD servers by exporting
their local \texttt{/dev/random} character devices. These devices are
all imported by a single client running some from on RNG aggregator
that combines the relatively slow output form the local RNGs into a
single high speed RNG output. We can effectively pool the entropy
available on numerous hosts to drive a single high performance RNG
system.

@@

Network Character Devices provide an extension of the standard device
driver file interface across the network. They allow for device
capabilities to be exported from where they are not needed and imported
where they are. By separating device usage from the device physical
location and host, the NCD system provides a generalized, transparent,
means for abstracting device functionality across a network.

>>>>

\usepackage[text={6.5in, 9in}, centering]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{listings}
\usepackage{hyperref}

@@

\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

@@

We have designed and implemented Network Character Devices (NCD) to
enable a new approach to using and sharing devices. Today, networking 
technology is becoming faster and more reliable at incredible rates. 
Functionality that has been traditionally believed to only be 
feasible at the local level can now be pushed to remote
machines. Our project aims to allow 
a character device present on a server to be exported to
a client over a network connection. Many systems use character device
interfaces, from Web cams, mice and monitors to USB flash drives and 
external hard drives. NCD works by abstracting the character device interface 
over IP. In Unix-like operating system, the character device interface
is a file interface. While designed and optimized for
character devices, this generality actually allows the NCD system to export 
any file. Being able to dynamically export devices over network
connections suggests a new cloud-like paradigm of device usage where 
locality is required only for reduced latency, which may not be a
concern for certain devices.

When a device is exported via the NCD system we would like it to give the
appearance of having essentially ``moved'' the device to the client 
by forbidding user level accesses on the server.
Acquiring a remote device will be
done through an advertisement protocol that exists in user space so
that details may be supplied about the server and how the connection
should be made (e.g. UDP vs TCP).  The client-side NCD's functionality
exists in kernel space so as to maintain transparency to
user level operations. Once imported, a device should be accessed just
like a normal local device.  This is achieved by the NCD kernel module
intercepting i/o calls and sending them out over the network to the
server. On the server, a userspace daemon accesses the physical
device to emulate the client's activity. The overall structure of the
NCD system is present in Figure \ref{fig:system-full}.

@@

Currently we have achieved a successful implementation of the low
level functionality for NCD. The client uses a kernel module written
for Linux 2.6.x+ while the server uses a userspace daemon. NCD is
still rough and in need of further development, as it tends to be
crash prone if not treated with care. The groundwork has been laid for
exporting any device, however many device subsystems access their
devices by avoiding the character device interface and these must be
dealt with on an individual basis. Section \ref{sec:futurework}
details some of these challenges and our plans for their resolution.

@@

NFS was the first implementation of generalized file sharing
over remote network connections.  However, NFS only supports normal
files and cannot interface directly with devices \cite{nfs-source}.

NBD (Network Block Device) was rolled into the Linux kernel starting
with version 2.1.101. It enables users to export block devices over the
network, allowing low level access to disks and other block devices.
However, NBD provides no
exclusive access guarantees and users must ascertain on their own
whether the remote drive is being used. It also mandates buffered
operation and is optimized for block, not character, device operation
\cite{nbd-source}.

CUSE (Character Device in User Space) is an extension of FUSE that
allows bringing a character device into a FUSE filesystem.  This allows
users to access the device directly and potentially export it using
user-level networking \cite{cuse-article}. We considered using CUSE in
our NCD implementation, but opted for a direct in-kernel implementation
instead since it is more standard.

USB/IP is a project to share USB devices over IP networks. This project is
very similar to NCD (as USB devices are character devices) but is
tailored specifically to allow sharing USB devices over a
network connection by intercepting USB specific data and encapsulating
it in IP \cite{usbip-source}. It does not support a more generalized
notion of general character device networking.

Plan 9 is a distributed operating system developed as a successor to
Unix by Bell Labs.  It implicitly allows sharing devices over a
network via file interfaces. However, Plan 9 does not currently see
wide deployment \cite{plan9-pike}. We hope to bring the NCD concept to
the more widely deployed Linux environment.

@@

The client is currently implemented entirely within an out-of-tree kernel
module. This is arguably not the most robust, nor secure, solution. See
Section \ref{sec:futurework} for details on splitting out some functionality
into a user-space helper.

@@

The server is a typical user-space daemon. It
currently only serves a single file. Whether or not the server should
handle multiple exports is still up for evaluation.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{proto-g.png}
  \caption{NCD operates over a simple request-reply protocol. Bulk data
    immediately follows formal messages, as required.}
  \label{fig:protocol}
\end{figure}

The protocol is implemented as a simple request-reply system over
\texttt{INET} sockets. All request types are encapsulated within a
single request structure (see Section \ref{sec:implementation}).
Similarly, there is only a single reply structure to account for all
reply types.

@@

Because there is only one type of packet for a request, and another for
a reply, (Listing \ref{lst:protocol}) there is a small amount of wasted
space on some file operations. We have done our best to reduce this
overhead as much as possible, making use of a cascade of nested unions
and structs.

@@

The client resides entirely within a kernel module \footnote{The merits
of including everything within a kernel module are discussed in section
\ref{sec:futurework}.}. A typical userspace program can be sloppy
with cleanup because all memory is freed on termination. In contrast, a
kernel module must be very careful because its lifetime is shorter than
the kernel itself. Therefore it is very import to correctly handle
errors and cleanup all memory in the \texttt{module\_init} and
\texttt{module\_exit} routines.

Some of the basic features of the NCD kernel module are listed in
Listing \ref{lst:client}. NCD dynamically requests a major and
minor number from the running kernel \footnote{Statically assigned
major/minor numbers are no longer required and are discouraged for new
modules.}. These numbers are stored in a \texttt{struct dev\_t}. Once
these numbers are assigned, the kernel will pass all file operations on
character devices with the NCD major/minor number to the NCD module.

@@

In the snippet given in Listing \ref{lst:server}, a \texttt{read()} file
operation is handled. A buffer must be allocated with a size
indicated by the original request. Next, the server can perform the
corresponding operation and send a reply message to the client. Finally,
the bulk data is sent to the client. All buffering is performed by the
underlying \texttt{SOCK\_STREAM} and TCP protocols.

@@

A lot of magic is performed on character devices in userspace by
the Linux \texttt{udev} system.
Currently we are using \texttt{udev} to manage the
permissions of imported devices (Listing \ref{lst:udevrule}).

@@

For more functionality, imported character devices must have certain
attributes such as those for the keyboard in Listing \ref{lst:udevattr}.
Unfortunately setting them retroactively does not convince all programs
of a particular devices type. Therefore interfacing with specific kernel
subsystems such as \texttt{input} and \texttt{usb} will be required in
the future for enhanced functionality.

@@

This set of core functionality provides a working proof-of-concept
Network Character Device system. This system fully supports the
export of a single character device from server to client over the
network and demonstrates the basic use and utility of such a system.

@@

Most modern Operating Systems, including Linux, assume that any device
physically connected to the system is available for the system to
use. As mentioned previously, Network Character Devices break down this
assumption. Thus, one of the challenges involved in implementing an NCD
system is to remove the OS's assumption that any connected device is
available for local use. Modern-day plug-and-play auto-configuration
systems make this task even more difficult, and the means for breaking
down the availability assumption tend to be unique to a given
OS. Effectively removing this assumption is one of the larger,
still-open, NCD challenges.

@@

\begin{packed_item}
\item Add multi-device, multi-server support
\item Add kernel sub-system handler layer (HID, etc)
\item Add device protection and removal capabilities to server system
\item Add advertisement protocol and unreliable data protocol option
\item Separate networking and memory-intensive code from client kernel module
\end{packed_item}

The current NCD system only supports a single device per server and per
client. To be truly useful and to enable the construction of the
kinds of system we envision NCD being used for (Section
\ref{sec:potentialuses}), we must at least
expand the client-side
system to support importing an arbitrary number of devices from one
or more servers. Ideally, servers should also be capable of serving
multiple devices, but this is of less importance than multi-device
client handling.

@@

Currently, all NCD traffic is sent via TCP. While control messages
should always use TCP to insure system state and stability, there are
device types who generate data more appropriate for UDP operation (web
cams, etc). In the future, we hope to allow the user to specify
whether the data traffic for a given device will be delivered via TCP
or whether it can be delivered using UDP.

@@

Finally, the current client-side NCD implementation performs both direct
network access and memory-intensive processing directly in the kernel
code. This approach was taken in our initial implementation to control
complexity.
For security reasons, it is desirable to move the direct
network communication from the kernel code to user-space code where
the repercussions of a security breach as less extreme.
For performance reasons, it is desirable to move memory
intensive operations from the kernel code (where they are limited to
the available kernel memory space) to user-space code where the memory
availability is far greater and the use restrictions more relaxed. To
solve these issues, we
foresee building a client-daemon that will handle the bulk to the
client-side operations and that will, in turn, communicate directly with
a more minimal client kernel module only when necessary.

@@

KVM (Keyboard-Video-Mouse) systems are commonly used in data centers
and other areas where a user wishes to use a single keyboard, video,
and mouse to control multiple machines. KVM systems range from simple KVM
switches that allow the user to select which local machine is
physically connected to his or her i/o devices to
networked KVM systems that allow the user to use a
local KVM setup to interface with a computer located elsewhere on the
network. The latter class of networked KVM devices traditionally require
expensive propriety hardware.

@@

Another example of the novel systems that can be built using Network
Charter Devices is a distributed random number generator. Linux has a
fairly robust random number generator built into it and accessible as
a Character Device at \texttt{/dev/random}. This RNG relies on the
local entropy pool (generally filled by human interaction, interrupt
timings, and other high entropy events) to generate random
numbers. Thus, the speed at which random numbers can be generated is
limited on any single system to the locally available
entropy. This can severely hinder the
performance of systems that require random numbers\footnote{Linux also
provides the \texttt{/dev/urandom} RNG which is a pseudorandom number
generator seeded off the true
\texttt{/dev/random}. \texttt{/dev/urandom} trades ``randomness'' in
favor of speed. It is thus handy when one just needs to quickly
generate pseudorandom numbers, but is not appropriate for cryptography or
other secure random system.}.

@@

Figure \ref{fig:rng} shows how such a system might work. In this
distributed RNG system, multiple hosts act as NCD servers by exporting
their local \texttt{/dev/random} character devices. These devices are
all imported by a single client running some from on RNG aggregator
that combines the relatively slow output from the local RNGs into a
single high speed RNG output. We can effectively pool the entropy
available on numerous hosts to drive a single high performance RNG
system.

@@

Network Character Devices provide an extension of the standard device
driver file interface across the network. They allow for device
capabilities to be exported from where they are not needed and imported
where they are. By separating device usage from the device's physical
location and host, the NCD system provides a generalized, transparent,
means for abstracting device functionality across a network.
